### 2017-03-22 paper/frag_implementation.tex
Start of implementation

\section{Implementation}

% smoother transition once we have everything

% Section intro:
% 1. Expectations (what hosts are there, and which ones run our code?)
% 2. Overview of the distinct pieces of software (client kernel, client daemon,
%    detour daemon)
% Subsection - Two Ways of Detours
% Subsection - Client kernel
% Subsection - Client daemon
% Subsection - Detour daemon

The implementation of this project is based on the following expectations.
First, the server must support Multipath TCP. However, beyond this, no other
modifications should be required of the server. Second, the client and detour
point are assumed to be ``active'' participants in the collective, and so they
may run whatever software is necessary to participate.

The client must run a modified kernel, which supports creating MPTCP subflows
across detours. In addition, the client must also run a daemon which discovers
and negotiates with detour points, finally reporting available ones to the
kernel. These concerns remain separate for several reasons. First, communication
with detour points requires communication over the Internet. While it is
possible to do user-level networking within the kernel, it is much simpler and
more future-proof to use the userspace socket API. Second, one type of detour
requires starting an OpenVPN process, an operation which should only be done
from userspace. Finally, userspace tools may be more easily configured and can
perform a more broad spectrum of computations, such as floating point math.

The detour need not support Multipath TCP. However, it must run one of two
possible daemons, depending on which type of detour it will be hosting.

\subsection{Detour Types}

In this project, we implement two different mechanisms for tunneling MPTCP
subflows across a detour.

\subsubsection{OpenVPN Tunneling}

The first mechanism involves the open-source tool OpenVPN. Detours run a
specially configured OpenVPN server, and clients connect. Unlike typical VPN
configurations, the client does not route all traffic through the VPN. Instead,
the client simply uses the VPN as a second network device. The client can then
use the VPN network device just as it would use any other physical device.

Traffic that goes through the VPN must undergo network address translation upon
reaching the detour point, since it functionally acts as a router.

\subsubsection{NAT Tunneling}

The second approach directly leverages NAT, without using OpenVPN for tunneling.
Since we only want to forward certain connections, we don't need the full power
of OpenVPN, and thus we can circumvent the overhead that comes with
encapsulation of datagrams.

The Linux kernel contains a robust packet filtering framework, known as
Netfilter. This framework includes utilities for performing network address
translation (NAT). This framework is typically what is used by consumer-grade
routers. Instead of encapsulation, the NAT tunneling approach directly addresses
the IP packets to the detour. The detour rewrites the source and destination
addresses using Netfilter for connection tracking.